Identification of Direct Cherenkov Pixels using Boosted Decision Trees

### 2016-06-21 report/draft1.tex
Backup 21/06/16

\documentclass{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.0in}
\usepackage{booktabs}
\usepackage{textcomp}
\begin{document}
\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

@@

Above threshold, Cherenkov Emission is determined by the Frank Tamm formula!!!!! A full simulation of the LPD was undertaken. Under the assumption of constant magnetic permittivity and a zenith angle of $90\textdegree$ we can use the approximate form \[ N_{photons} \approx (37000 \times Z^{2} \times \Delta_{h} \times \Delta_{E})\] where $\Delta_{h}$ is the vertical distance travelled in meters and $\Delta_{E}$ is the emission energy range in eV.

@@

Thus the upper atmosphere emission contributes to the inner LDF, while the lower atmosphere emission contributes to the outer LDF. We can also see ground emission radius as a function of height in \ref{fig:generalenergy}. We find that the high-radius emission (occurring near the first interaction region) varies little between different high energies. We deem this to be \textquoteleft Saturated Emission\textquoteright.

@@

\subsection{First Interaction Height}
Cosmic Rays survival in from the top of the atmosphere follows an exponential decay with the number of 'interaction lengths' passed. The interaction length is dependent on the interaction cross section, which increases with density. Thus the number of interaction lengths increases exponentially as height decreases. The resultant decays occur most often at a height of $h \approx 40 \pm 10$ km , as seen in \ref{fig:generalheight}.

\subsection{Atmospheric Absorption}
The Cherenkov Light, mostly emitted in the visible blue part of the EM spectrum, experiences relatively little atmospheric absorption. The major of Rayleigh scattering-based atmospheric absorption occurs in the lower part of the troposphere, and thus the atmospheric absorption is almost independent of emission up to first interaction height.

@@

\subsection{Image Simulation}
The full simulation of air showers was performed using the CORSIKA package, with a standard atmospheric profile derived from measurements conducted at the HESS site in Namibia. In total, 5000 training events and a further 10000 testing events were simulated. The simulated particles were $Fe^{56}$, within the Energy Range of $35-135$ TeV and a spectrum $\phi(E) \propto E^{-2.7}$. For each set of simulated event, 4 unique random number seeds were used to generate the shower. An altitude of 1800m was assumed, again corresponding to the HESS site. The simulated zenith angle ranged from $0\textdegree < \theta < 2 \textdegree$, while the simulated azimuth angle ranges from -2\textdegree $ < \phi < 2 \textdegree$. The four smaller HESS-phase-1 telescopes were arranged in a cross along the x/y axis with the larger HESS-phase-2 \textquoteleft CT5' telescope placed at the center. The length of each cross arm was $85m$. The simulated target region of the cores was chosen to be a square centered on CT5, with each 300m-long side bisecting the x/y axis.Due to hardware differences between CT5 and the original HESS 1 telescopes, we analyse HESS1 and HESS2 images separately. To determine the true class of each pixel, a simulation was initially run with an energy cut of 10 PeV on all muons and electrons. Because this cut exceeded the primary particle energy, neither daughter muons and electrons, nor the photons they would have emitted, were simulated. Thus the hadronic Cherenkov Light from the primary particle and daughter fragments, but not the EAS light, was present in the camera image. A second identical \textquoteleft EAS Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete EAS image including identical DC light.

@@

@@

\section{HESS-type Event Reconstruction}
Having obtained a parameterisation of the LPD, and and assoiated error in the case of a HESS-type array, the Charge reconstruction Technique can be applied to a simpified Monte Carlo simulation of the HESS array, developed as part of this analysis. Conducting a Monte Carlo simulation using the Gaussian-smeared LPD, we can measure the effectiveness of the LPD-reconstruction technique.

@@

To accurately model the Cosmic Rays, the Energy Power Law must be simulated. To generate the random Energy value En(R) following the power law, a uniformly distributed random number R is generated. The value of R represents a random fraction of the total simulated flux $ F(R) = R \times F(35-135TeV) $, and can thus range from 0-1. Having determined the Integrated Flux corresponding to the random number, the Energy corresponding to this flux can be calculated. The energy E is defined as the lower bound of the integral which would produce the chosen integrated flux F. Thus R=0 corresponds to an Energy of 135TeV yielding no Integrated Flux, while R=1 would correspond to an Energy of 35TeV and the full Integrated Flux.

@@

\subsection{First Interaction and Energy Saturation}
Using the theoretical LPD found from the Frank-Tamm formula, we see that this LPD will have a very abrupt discontinuity. However, after the first atmospheric simulation, secondary DC light from daughter nucleii will also be produced. it is not possible to effectively distinguish this secondary DC light, so it will contribute to the smearing seen in Figure \ref{fig:corsikalpd1}. The first Interaction Height will determine the radius at which the LPD for a given Iron Nucleus will deviate from the ideal case. By considering the distribution of first interaction heights, we can consider the variation of the LPD.

>>>>

\documentclass{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.5in}
\usepackage{gensymb}
\usepackage{booktabs}
\begin{document}
\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

@@

Above threshold, Cherenkov Emission is determined by the Frank Tamm formula \cite{ppreview}.  Under the assumption of constant magnetic permittivity and a zenith angle of $90\degree$, and with  the vertical distance travelled in meters $\Delta_{h}$ and the emission energy range in eV $\Delta_{E}$, we can find an approximate Cherenkov Emission formula.
\[ \frac{d^{2}N_{photons}}{dEdh} = \frac{\alpha Z^{2}}{\hbar c} \sin^{2} \theta_{c}(E) \approx 370 Z^{2} \sin^{2} \theta_{c}(E)\, cm^{-1}eV^{-1} \]

\[ N_{photons} \approx (37000 \times Z^{2} \times \Delta_{h} \times \Delta_{E}(E))\, m^{-1}eV^{-1}\]

A full simulation of the LPD was undertaken, with the assumption that $\Delta_{E} \approx 0.7$ independent of Energy. Although this is certainly an oversimplification of the true picture, it will allow us to understand the variation of the LPD with height. A more rigorous LPD parameterisation will be found later to account for the dependence of $\Delta_{E}$ on the Energy.

@@

Thus the upper atmosphere emission contributes to the inner LDF, while the lower atmosphere emission contributes to the outer LDF. As the cosmic ray travels down through the atmosphere, Cherenkov Emission continuously illuminates new areas on the ground that were previously unlit, without re-illuminating inner regions. This is extremely useful, because it means that for all distances up to the maximum radius, the LPD will be completely independent of the emission height. One full parameterisation of the LPD will thus be valid for all heights, provided the maximum radius is known.

We can also see ground emission radius as a function of height in \ref{fig:generalenergy}. We find that the high-radius emission (occurring near the first interaction region) varies little between different high energies. We deem this to be \textquoteleft Saturated Emission\textquoteright.

@@

@@

\subsection{Image Simulation}
The full simulation of air showers was performed using the CORSIKA package, with a standard atmospheric profile derived from measurements conducted at the HESS site in Namibia. In total, 5000 training events and a further 10000 testing events were simulated. The simulated particles were $Fe^{56}$, within the Energy Range of $35-135$ TeV and a spectrum $\phi(E) \propto E^{-2.7}$. For each set of simulated event, 4 unique random number seeds were used to generate the shower. An altitude of 1800m was assumed, again corresponding to the HESS site. The simulated zenith angle ranged from $0\degree < \theta < 2 \degree$, while the simulated azimuth angle ranges from $-2\degree < \phi < 2 \degree$. The four smaller HESS-phase-1 telescopes were arranged in a cross along the x/y axis with the larger HESS-phase-2 \textquoteleft CT5' telescope placed at the center. The length of each cross arm was $85m$. The simulated target region of the cores was chosen to be a square centered on CT5, with each 300m-long side bisecting the x/y axis.Due to hardware differences between CT5 and the original HESS 1 telescopes, we analyse HESS1 and HESS2 images separately. To determine the true class of each pixel, a simulation was initially run with an energy cut of 10 PeV on all muons and electrons. Because this cut exceeded the primary particle energy, neither daughter muons and electrons, nor the photons they would have emitted, were simulated. Thus the hadronic Cherenkov Light from the primary particle and daughter fragments, but not the EAS light, was present in the camera image. A second identical \textquoteleft EAS Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete EAS image including identical DC light.

@@

\section{HESS-type Event Reconstruction}

@@

\subsection{Other Stuff Like}
Having obtained a parameterisation of the LPD, and and assoiated error in the case of a HESS-type array, the Charge reconstruction Technique can be applied to a simpified Monte Carlo simulation of the HESS array, developed as part of this analysis. Conducting a Monte Carlo simulation using the Gaussian-smeared LPD, we can measure the effectiveness of the LPD-reconstruction technique.

@@

To accurately model the Cosmic Rays, the Energy Power Law must be simulated. To generate the random Energy value En(R) following the power law, a uniformly distributed random number R is generated. The value of R represents a random fraction of the total simulated flux $ F(R) = R \times F(35-135TeV) $, and can thus range from 0-1. Having determined the Integrated Flux corresponding to the random number, the Energy corresponding to this flux can be calculated. The energy E is defined as the lower bound of the integral which would produce the chosen integrated flux F. Thus R=0 corresponds to an Energy of 135TeV yielding no Integrated Flux, while R=1 would correspond to an Energy of 35TeV and the full Integrated Flux. THus the simulated Cosmic Rays will obey the power law, with high-energy events being realistically suppressed.

@@

\subsection{First Interaction Height}
Cosmic Rays survival in from the top of the atmosphere follows an exponential decay with the number of 'interaction lengths' passed.

The mean free path $\ell$ of a Cosmic Ray is a function of atmospheric cross section and number density, which are themselves functions of  height, so that $\ell(h) = \frac{1}{\sigma(h) n(h)}$. It is instead easier to consider a new variable x, which we name the \textquoteleft Interaction Distance', so that the survival probability of a cosmic ray follows a simple exponential decay with the interaction distance passed. We thus define one Interaction Length, $x_{0}$ such that a cosmic ray traveling through one interaction length will have a non-interaction probability of $\frac{1}{e}$, and in general $P_{survival} \propto e^{\frac{-x}{x_{0}}}$ The geometric distance corresponding to one interaction length will vary as a function of height.

The Interaction Length is inversely proportional to the interaction cross section, which is itself proportional to local atmospheric number density, so that $x_{0} \propto \frac{1}{n(h)}$. Thus, because the number density of the atmosphere increases with decreasing height, the local interaction length will be geometrically shortened as the height of the cosmic ray decreases. If we consider the Integrated Interaction Distance as the total number of interaction lengths passed by a Cosmic Ray from the top of the atmosphere to a given height h, we see in Figure \ref{fig:generalheight} that the total integrated interaction distance increases exponentially with increasing height. We use the standard HESS-based data tables also found in the CORSIKA software, containing information regarding the integrated interaction distance and refractive index over a range of heights. As a function of x, the corresponding local interaction rate $\Gamma_{interact}(x)$ will be a product of the survival probability and the probability of interacting $P_{interact}(x)= -\frac{dP_{survival}(x)}{dx}$.

\[ \Gamma_{interact}(x) = P_{interact} \times N_{survived} =  - \frac{dP_{survival}(x)}{dx} \times (N \times  P_{survival}) \]

The fraction of surviving and interacting cosmic rays are shown in Figure \ref{fig:generalheight}, alongside the resultant distribution of interactions binned by height h rather than interaction length x. The height of each bin can be considered an average local value for $\Gamma_{interact}(h)$, and it is clear that the function $\Gamma_{interact}(h)$ peaks at a height of around 40km. This is thus the height region where most of our cosmic rays will interact with the atmosphere.

To simulate the First Interaction Height, we assign a uniform random number R to each event, corresponding to the $P_{survival}$ at which the cosmic ray will interact. R can be interpreted fraction of events that will have interacted before the simulated cosmic ray does. It ranges from R=0 for interaction at the top of the Atmosphere to R=1 for an event reaching the ground. Having determined the fraction of events, we assume the survival probability is normalised. As with all exponential distributions, we see that the expectation value is equal to one Interaction Length.

\[ P_{survival}(x) = k e^{\frac{-x}{x_{0}}} \Longrightarrow \int_{0}^{\infty} P_{survival}(x) dx = k x_{0} [ e^{0} - e ^ {- \infty}] = 1 \Longrightarrow k = \frac{1}{x_{0}}\]
\[P_{half}(x) =\int_{0}^{x_{median}} k e^{\frac{-x}{x_{0}}} dx =[1 -e^{\frac{-x_{median}}{x_{0}}} ]= \frac{1}{2} \Longrightarrow {x_{median}} = x_{0} \ln (2)\]

\[ <x> =  \int_{0}^{\infty} P_{survival}(x) dx = x_{0} = \frac{x_{median}}{\ln (2)} \]

Thus, we can simply calculate $x_{0}$ as the expectation value of interaction length. This parameter is free-floating, and can only be determined through experimental study of Cosmic Rays. Experimental results indicate an approximate cross section of $\sigma_{Fe-air} = 2000mb$ and a corresponding thus $x_{0} \approx 12 g cm^{-2}$ \cite{Montanus2013}.

As with energy, we wish to simulate a realistic spectrum for the first interaction heights. We must thus convert the uniform random number R to an interaction distance X. Using the standard atmospheric tables, the corresponding height value can be found using exponential interpolation.

\[ R = \int_{0}^{X} k e^{\frac{-x}{x_{0}}} dx = [1 -e^{\frac{-X}{x_{0}}}] \Longrightarrow X = -x_{0} \ln (1 - R) \]

But you lose the tail..

As cited by the HESS study, a typical first interaction height is roughly 30km for a Cherenkov-Emitting event \cite{hess07}. With the assumed value of $x_{0}$, a simulation was conducted to determine the expected interaction heights for Cosmic Rays. It is found that the mean first interaction height of 36.9km and a median first interaction height of 35.2km for all cosmic rays. However, when Cherenkov-Emitting events are considered, we have both a median and mean first interaction height of 30.3km, which clearly agrees with our expectations.

Another parameter exponentially interpolated using the data tables is the optical depth to the surface, $\tau$. The Cherenkov Light, mostly emitted in the visible blue part of the EM spectrum, experiences relatively little atmospheric absorption. The major of Rayleigh scattering-based atmospheric absorption occurs in the lower part of the troposphere, and thus the atmospheric absorption is almost independent of emission up to first interaction height. For a given emitted intensity $I_{0}$ we find that the corresponding I received on the ground decays exponentially with optical depth, so that $\frac{I}{I_{0}}=e^{-\tau}$ with $\tau=\int_{0}^{l}ds$.

\subsection{First Interaction and Energy Saturation}
Using the theoretical LPD found from the Frank-Tamm formula, we see that this LPD will have a very abrupt discontinuity.
